---
title: "FM: - Week2 - Assignment"
author: "Thanasit C."
date: "2024-01-24"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week2: Assignment

### 1. Packages
install.packages("ggplot2")\
install.packages("tidyverse")\
install.packages("readxl")\
install.packages("dplyr")\
install.packages("ggpubr")\


### 2. Libraries
```{r packages, warning=FALSE, results='hide', message=FALSE}
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
library(ggpubr)
```


### 3. Set working directory
```{r}
setwd("/Users/sieng/Documents/Study/MS.Business Analytics/SPRING 2024/Fundamental of Machine Learning/Assignment/Assignment 1")
```


### 4. Import data [Top 50 stocks in The Stock Exchange of Thailand: SET50]
source: https://www.set.or.th/en/market/index/set50/overview
```{r}
set50df <- read_excel("FML - Assignment_1 - dataset - SET50.xlsx") %>% as.data.frame()
str(set50df)
```

### 5. Data manipulation

#### 5.1 Change column name
It is easier not to use "%" or "()" as a column name
```{r}
colnames(set50df)[7] <- "DailyRetrun"
colnames(set50df)[10] <- "Volume"
colnames(set50df)[11] <- "Value"
```

#### 5.2 Ranking
Create new column (variable) by ranking column $DailyReturn. Then change its class from "Numeric" to "Factor with ordered"\
Then I order by its rank and find that "TRUE" is perform the best.  
```{r}
set50df$Ranking <- rank(-set50df$DailyRetrun)
class(set50df$Ranking)

set50df$Ranking <- factor(set50df$Ranking, order = TRUE)
class(set50df$Ranking)

set50rankdf <- data.frame(set50df$Symbol, set50df$Ranking)
colnames(set50rankdf)[1] <- "Symbol"
colnames(set50rankdf)[2] <- "Ranking"
set50rankdf[order(set50rankdf$Ranking),]
```

#### 5.3 Convert1 
To convert class of $Industrial column from "Character" to "Factor" variable\
Initial class of $Industry is "Character". Then convert to "Factor"
```{r}
class(set50df$Industry)
set50df$Industry <- as.factor(set50df$Industry)
class(set50df$Industry)
```

#### 5.4 Convert2
To convert $Value unit from "'000 Bahts" to "millions baht"
```{r}
set50df$Value <- set50df$Value/1000 
```


#### 5.5 Create new table 
To explore more about the relationship between $DailyReturn and $Industry
```{r}
set50Inddf <- set50df[c(12,7,11)]
set50Inddf2 <- set50df %>% 
                 group_by(Industry) %>%
                 count(Industry)
set50Inddf3 <- set50Inddf %>%
                  group_by(Industry) %>%
                  dplyr::summarise(AvgValue = mean(Value), AvgReturn = mean(DailyRetrun))
set50Inddf4 <- merge(set50Inddf3, set50Inddf2, by.x = "Industry")
colnames(set50Inddf4)[4] <- "NumberOfStocks"
str(set50Inddf4)
```


### 6. Descriptive

#### 6.1 Data summary
```{r}
summary(set50df)
```

```{r}
summary(set50Inddf4)
```

#### 6.2 Statistics
For $DailyReturn and Value.
```{r}
statistics <-  c("mean", "max", "min", "range", "StdDev", "variance", "correlation")
DailyReturn <- c(mean(set50df$DailyRetrun), 
                 max(set50df$DailyRetrun), 
                 min(set50df$DailyRetrun), 
                 max(set50df$DailyRetrun) - min(set50df$DailyRetrun),
                 sd(set50df$DailyRetrun),
                 var(set50df$DailyRetrun),
                 cor(set50df$DailyRetrun, set50df$Value))
Value <- c(mean(set50df$Value), 
                 max(set50df$Value), 
                 min(set50df$Value), 
                 max(set50df$Value) - min(set50df$Value),
                 sd(set50df$Value),
                 var(set50df$Value),
                 cor(set50df$Value, set50df$DailyRetrun))
statdf <- data.frame(statistics, DailyReturn, Value)
statdf$DailyReturn <- round(statdf$DailyReturn, digits = 2)
statdf$Value <- round(as.numeric(format(statdf$Value, scientific = FALSE)), digits = 2)
statdf
```


#### 6.3 Visualization
First, I create [Scatter plot] between stocks daily return and trade value.\
Then I plot [Histogram] to see the distribution of the stock return.\
Next, create a [Column chart] to explore diversity of stocks in SET50.\
Lastly, another [Scatter plot with linear regression] is created to investigate Industry daily return and Industry trade value.

```{r, warning=FALSE}
set50df %>% ggplot(mapping = aes(x = Value, y = DailyRetrun, label = Symbol)) +
              geom_point() + 
              geom_text() +
              geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
              labs(title = ("The Stock Exchange of Thailand top 50 stocks"), 
                   subtitle = ("Scatter plot between Stocks Daily return and Trade Value"), 
                   caption = ("As of Jan 24, 2024")) +
              xlab("Value (millions)") +
              ylab("Daily return (%)")
  
set50df %>% ggplot(data = set50df, mapping = aes(x = DailyRetrun)) +
              geom_histogram(binwidth = 0.25) +
              geom_vline(mapping = aes(xintercept = mean(DailyRetrun)), 
                         color = "blue", 
                         linetype = "dashed", 
                         size = 1) +
              labs(title = ("The Stock Exchange of Thailand top 50 stocks"), 
                   subtitle = ("Daily Return Distribution"), 
                   caption = "As of Jan 24, 2024") +
              xlab("Count (n)") +
              ylab("Daily Return (%)")  
  
set50Inddf4 %>% ggplot(mapping=aes(x=Industry, y=NumberOfStocks, label = NumberOfStocks)) +
                  geom_col(fill = "lightblue") +
                  geom_text() +
                  labs(title = ("The Stock Exchange of Thailand top 50 stocks"), 
                       subtitle = ("Industry diversification"), 
                       caption = "As of Jan 24, 2024") +
                  ylab("Number of Stocks")

set50Inddf4 %>% ggplot(mapping = aes(x = AvgValue, y = AvgReturn, label = Industry)) +
                  geom_point() +
                  geom_text() +
                  geom_smooth(method = "lm", formula = y~x) +
                  stat_cor(label.y = 0.75,
                           mapping = aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"))) +
                  labs(title = ("The Stock Exchange of Thailand top 50 stocks"), 
                       subtitle = ("Relationship between Indusrty and Daily Return"), 
                       caption = "As of Jan 24, 2024") +
                  xlab("Value (millions)") +
                  ylab("Daily Return (%)")

```




